DeepSphere: Efficient spherical Convolutional Neural Network with HEALPix sampling for cosmological applications

### 2018-11-02 scnn.tex
coma
typo

Following~\cite{shuman2013emerging}, the normalized graph Laplacian,
defined as $\L = \I - \D^{-1/2} \W \D^{-1/2}$, is a second order differential operator
that can be used to define a Fourier basis on the graph. Here $\D$ is the diagonal
matrix where $\D_{ii} = \b{d}_i$ and $\b{d}_i = \sum_j \W_{ij}$ is the weighted degree of node $v_i$. By construction, the Laplacian is symmetric positive
semi-definite and hence can be decomposed as $\L = \U \bLambda \U\trans$, where $\U = [\b u_1, \ldots, \b u_{N_{pix}}]$ is an
orthonormal matrix of eigenvectors and $\bLambda$ is a diagonal matrix of
eigenvalues. The graph Fourier basis is defined as the Laplacian eigenvectors, motivated by the fact that a Fourier basis should diagonalize the Laplacian operator.
The graph Fourier transform of a signal $\f \in \R^{N_{pix}}$ is simply its projection on the eigenvectors given by
$\hat{\f} = \mathcal{F}_\G \{\f\} = \U\trans \f$. It follows that the inverse graph Fourier transform reads $\mathcal{F}^{-1}_\G \{\hat{\f}\} = \U\hat{\f} = \U \U\trans \f = \f$.
Note that the Fourier modes are ordered in the increasing order of the Laplacian eigenvalues $\bLambda$, which can be interpreted as squared frequencies.
Indeed,
\begin{equation*}
	\bLambda_{ii} = \b u_i\trans \L \b u_i = \sum_{(v_j, v_k) \in \E} \frac{\W_{jk}}{\sqrt{\b d_j \b d_k}} (\U_{ji} - \U_{ki})^2
\end{equation*}
is a measure of the variation of the eigenvector $\b u_i$ is on the graph defined by the Laplacian $\L$.

Figure \figref{graph_harmonics} shows the Fourier modes of a HEALPix graph, created using the graph construction described above.
The graph Fourier modes resemble the spherical harmonics.
That is a strong hint that the graph is able to capture the spherical properties of the HEALPix sampling.
This topic is further discussed in \ref{sec:comparison_spherical_harmonics}.

@@

We present DeepSphere, a novel convolutional neural network defined on the sphere with HEALPix sampling, designed for the analysis of the cosmological data.
We publish the code as a small and easy-to-use python package.
The code was designed so that DeepSphere can easily be used for typical machine learning tasks, such as classification and regression, both dense and global.
The core of our method is the use of a graph to represent the discretised sphere.
This allows us to leverage the advantages of the graph convolution.
It is both efficient, with a complexity of $\bO(N_{pix})$, and flexible, which allows DeepSphere to efficiently work on parts of the sphere.
The spherical properties of the domain are well captured: the graph Fourier modes are close to the spherical harmonics.
Filters are restricted to be radial for the convolution operation to be equivariant to rotation.
DeepSphere can then be made either equivariant or invariant to rotation.
Equivariance is not perfect as small imprecisions due to the sampling cause the action of a graph filter to slightly depend on the location.
However, we do not expect that to cause problems for practical application

@@

Since the convergence map characterises the dark matter lenses at different redshifts, one has to first construct a lightcone. The lightcone represents the volume of the universe around the observer, where the radius represents the distance the light from distant galaxies at different times and locations has to travel to the observer. Therefore, the particles outputted at different redshifts (time-steps) of the simulation are concentrically arranged in shells of constant redshift around the observer (each redshift $z$ corresponds to a comoving distance away from the observer $\chi(z)$), which is located at redshift $z=0$ (see Figure \ref{fig:sketch_lightcone_simulation}). The shells are then weighted by a weak lensing specific weight $W_b$ and projected onto the sphere for each pixel $\theta_\mathrm{pix}$, resulting in a convergence map
\begin{equation}
\kappa (\theta_\mathrm{pix}) \sim \sum_b W_b \frac{V_\mathrm{sim}}{N_\mathrm{part}^\mathrm{sim}} \frac{n_p (\theta_\mathrm{pix}, \Delta \chi_b)}{\mathcal{D}^2 (z_b)}  \, ,
\end{equation}
where $n_p$ represents the number of particles in a specific pixel on the shell $\Delta \chi_b$. The convergence map depends on the cosmological parameters.

>>>>

Following~\cite{shuman2013emerging}, the normalized graph Laplacian,
defined as $\L = \I - \D^{-1/2} \W \D^{-1/2}$, is a second order differential operator
that can be used to define a Fourier basis on the graph. Here $\D$ is the diagonal
matrix where $\D_{ii} = \b{d}_i$ and $\b{d}_i = \sum_j \W_{ij}$ is the weighted degree of node $v_i$. By construction, the Laplacian is symmetric positive
semi-definite and hence can be decomposed as $\L = \U \bLambda \U\trans$, where $\U = [\b u_1, \ldots, \b u_{N_{pix}}]$ is an
orthonormal matrix of eigenvectors and $\bLambda$ is a diagonal matrix of
eigenvalues. The graph Fourier basis is defined as the Laplacian eigenvectors, motivated by the fact that a Fourier basis should diagonalize the Laplacian operator.
The graph Fourier transform of a signal $\f \in \R^{N_{pix}}$ is simply its projection on the eigenvectors given by
$\hat{\f} = \mathcal{F}_\G \{\f\} = \U\trans \f$. It follows that the inverse graph Fourier transform reads $\mathcal{F}^{-1}_\G \{\hat{\f}\} = \U\hat{\f} = \U \U\trans \f = \f$.
Note that the Fourier modes are ordered in the increasing order of the Laplacian eigenvalues $\bLambda$, which can be interpreted as squared frequencies.
Indeed,
\begin{equation*}
	\bLambda_{ii} = \b u_i\trans \L \b u_i = \sum_{(v_j, v_k) \in \E} \frac{\W_{jk}}{\sqrt{\b d_j \b d_k}} (\U_{ji} - \U_{ki})^2
\end{equation*}
is a measure of the variation of the eigenvector $\b u_i$ on the graph defined by the Laplacian $\L$.

\figref{graph_harmonics} shows the Fourier modes of a HEALPix graph, created using the graph construction described above.
The graph Fourier modes resemble the spherical harmonics.
That is a strong hint that the graph is able to capture the spherical properties of the HEALPix sampling.
This topic is further discussed in \ref{sec:comparison_spherical_harmonics}.

@@

We present DeepSphere, a novel convolutional neural network defined on the sphere with HEALPix sampling, designed for the analysis of the cosmological data.
We publish the code as a small and easy-to-use python package.
The code was designed so that DeepSphere can easily be used for typical machine learning tasks, such as classification and regression, both dense and global.
The core of our method is the use of a graph to represent the discretised sphere.
This allows us to leverage the advantages of the graph convolution.
It is both efficient, with a complexity of $\bO(N_{pix})$, and flexible, which allows DeepSphere to efficiently work on parts of the sphere.
The spherical properties of the domain are well captured: the graph Fourier modes are close to the spherical harmonics.
Filters are restricted to be radial for the convolution operation to be equivariant to rotation.
DeepSphere can then be made either equivariant or invariant to rotation.
Equivariance is not perfect as small imprecisions due to the sampling cause the action of a graph filter to slightly depend on the location.
However, we do not expect that to cause problems for practical applications.

@@

Since the convergence map characterises the dark matter lenses at different redshifts, one has to first construct a lightcone. The lightcone represents the volume of the universe around the observer, where the radius represents the distance the light from distant galaxies at different times and locations has to travel to the observer. Therefore, the particles outputted at different redshifts (time-steps) of the simulation are concentrically arranged in shells of constant redshift around the observer (each redshift $z$ corresponds to a comoving distance away from the observer $\chi(z)$), which is located at redshift $z=0$ (see \figref{sketch_lightcone_simulation}). The shells are then weighted by a weak lensing specific weight $W_b$ and projected onto the sphere for each pixel $\theta_\mathrm{pix}$, resulting in a convergence map
\begin{equation}
\kappa (\theta_\mathrm{pix}) \sim \sum_b W_b \frac{V_\mathrm{sim}}{N_\mathrm{part}^\mathrm{sim}} \frac{n_p (\theta_\mathrm{pix}, \Delta \chi_b)}{\mathcal{D}^2 (z_b)}  \, ,
\end{equation}
where $n_p$ represents the number of particles in a specific pixel on the shell $\Delta \chi_b$. The convergence map depends on the cosmological parameters.
